draft_final_assignment

### 2022-11-30 draft_final_assingment.Rmd
Summary_final_assignment.Rmd

---
title: "draft_final_assignment"
author: "Andrew Wells"
date: "11/18/2022"
output: 
  html_document:
    toc: TRUE
    theme: cosmo
    toc_float: TRUE
---
# Research question

@@

-   While one might expect that the post 2016 and 2020 congresses will vary in their topics and policies. However, it is interesting to also focus on the period between 2018 and 2020: When both chambers had different majorities and needed to cooperate.

@@

```{r }
df <- read_csv("https://raw.githubusercontent.com/juka19/tad_assignment3/main/data/data_11_28.csv")
```

@@

```{r}
ggplot(df, aes(x = cosponsor_D_perc)) + 
  geom_histogram(aes(y=..density..), colour="black", fill="white") +
  geom_density(alpha=.1, fill="blue") +
  labs(title="Density of bill cosposor party",
       x ="Cosponsor party composition", y = "Density", 
       caption = "Numbers represent proportion of cosponsors from Democratic party, 
       so 0.0 represents bills that were fully Republican and 1.0 represents 
       bills that were fully Democrat.") +
  theme_minimal()
```

@@

## Creatung a dfm from the corpus

```{r }
dfmat <- corp %>%
  tokens(remove_punct = TRUE, remove_numbers = TRUE, remove_symbols = TRUE) %>%
  tokens_remove(patter = stopwords("en")) %>%
  tokens_replace(pattern = lexicon::hash_lemmas$token, replacement = lexicon::hash_lemmas$lemma) %>%
  tokens_wordstem() %>%
  tokens_remove(c("sec","bill","act", "section", "funds", "shall","must", "use", "author","fund","provid","program","requir","divis","titl","appropri","specifi")) %>%
  dfm()

#Do we also want to lemmatize and apply word stems??
# tokens_replace(pattern = lexicon::hash_lemmas$token, replacement = lexicon::hash_lemmas$lemma) %>%
# tokens_wordstem() %>%
```

## Most common words in the corpus

```{r }
dfmat %>% colSums() %>% sort(decreasing = TRUE) %>% head(20)
```

@@

## Wordcloud congress 115

```{r, wa }
dfmat_115 <- dfm_subset(dfmat, session == 115)

textplot_wordcloud(dfmat_115, max_words = 300)
```

## Wordcloud congress 116

```{r, wa }
dfmat_116 <- dfm_subset(dfmat, session == 116)

textplot_wordcloud(dfmat_116, max_words = 300)
```

## Wordcloud congress 117

```{r, wa }
dfmat_117 <- dfm_subset(dfmat, session == 117)

textplot_wordcloud(dfmat_117, max_words = 300)
```

@@

## Wordcloud congress parties 117

corp_117 <- df %>% filter(session == 117) %>% rename(text = summary) %>% corpus()
modelpart17 <- dfm(corp_117, remove = stopwords("english"), remove_numbers = TRUE, remove_punct = TRUE, groups = corp_117$party) %>% dfm_remove(c("sec","bill","act", "section", "funds", "shall","must", "used")) %>%
  dfm_trim(min_termfreq = 3)

mp17 <- textplot_wordcloud(modelpart17, comparison = TRUE, max_words = 300,
                   color = c("green","blue", "red"))

mp17

## Wordcloud comparing Congresses

```{r, wa }
dfmatCon <- dfm(corp, remove = stopwords("english"), remove_numbers = TRUE, remove_punct = TRUE, groups = corp$session) %>%dfm_remove(c("sec","bill","act", "section", "funds", "shall","must", "used")) %>%
  dfm_trim(min_termfreq = 3)

textplot_wordcloud(dfmatCon, comparison = TRUE, max_words = 300,
                   color = c("blue", "red"))
```

## Most common words (overall)

```{r, warning=FALSE}
#Overall most common words

tfreq <- dfmat %>%
  textstat_frequency() %>%
  head(20)

tfreq$feature <- factor(tfreq$feature, levels=tfreq$feature)

ggplot(tfreq, aes(x=frequency, y=feature)) +
geom_col() + labs(title="Feature (word) frequency",
        x ="Frequency", y = "Feature") +
  theme_minimal() 
```

## Most common words by party

```{r, warning=FALSE}
#Republicans
dfmat_r <- dfm_subset(dfmat, party == "Republican")

tfreq_r <- dfmat_r %>%
  textstat_frequency() %>%
  head(20)

tfreq_r$feature <- factor(tfreq_r$feature, levels=tfreq_r$feature)

ggplot(tfreq_r, aes(x=frequency, y=feature)) +
geom_col() + labs(title="Feature (word) frequency for Republican Bills",
        x ="Frequency", y = "Feature") +
  theme_minimal() 
```

```{r, warning=FALSE}
#Democrats
dfmat_d <- dfm_subset(dfmat, party == "Democrat")

tfreq_d <- dfmat_d %>%
  textstat_frequency() %>%
  head(20)

tfreq_d$feature <- factor(tfreq_d$feature, levels=tfreq_d$feature)

ggplot(tfreq_d, aes(x=frequency, y=feature)) +
geom_col() + labs(title="Feature (word) frequency for Democrat Bills",
        x ="Frequency", y = "Feature") +
  theme_minimal() 
```

```{r, warning=FALSE}
party_corp <- corpus_subset(corp, 
                            party %in% c("Democrat", "Republican"))

# Create a dfm grouped by party
party_dfm <- tokens(party_corp, remove_punct = TRUE, remove_numbers = TRUE, remove_symbols = TRUE) %>%
  tokens_remove(patter = stopwords("en")) %>%
  tokens_replace(pattern = lexicon::hash_lemmas$token, replacement = lexicon::hash_lemmas$lemma) %>%
  tokens_wordstem() %>%
  tokens_remove(c("sec","bill","act", "section", "funds", "shall","must", "use", "author", "may")) %>%
  tokens_group(groups = party) %>%
  dfm()

# Calculate keyness and determine Republican as target group
result_keyness <- textstat_keyness(party_dfm, target = "Republican")

# Plot estimated word keyness
textplot_keyness(result_keyness) 
```

## Most common words (by session)

```{r, warning=FALSE}
# Most common words by Congressional session 115
tfreq_115 <- dfmat_115 %>%
  textstat_frequency() %>%
  head(20)

tfreq_115$feature <- factor(tfreq_115$feature, levels=tfreq_115$feature)

ggplot(tfreq_115, aes(x=frequency, y=feature)) +
  geom_col() + labs(title="Feature (word) frequency for 115 session (2017 - 2018)",
        x ="Frequency", y = "Feature") +
  theme_minimal()
```

```{r, warning=FALSE}
# Most common words by Congressional session 116
tfreq_116 <- dfmat_116 %>%
  textstat_frequency() %>%
  head(20)

tfreq_116$feature <- factor(tfreq_116$feature, levels=tfreq_116$feature)

ggplot(tfreq_116, aes(x=frequency, y=feature)) +
  geom_col() + labs(title="Feature (word) frequency for 116 session (2019 - 2020)",
        x ="Frequency", y = "Feature") +
  theme_minimal()
```

```{r, warning=FALSE}
# Most common words by Congressional session 117
tfreq_117 <- dfmat_117 %>%
  textstat_frequency() %>%
  head(20)

tfreq_117$feature <- factor(tfreq_117$feature, levels=tfreq_117$feature)

ggplot(tfreq_117, aes(x=frequency, y=feature)) +
  geom_col() + labs(title="Feature (word) frequency for 117 session (2021 - 2022)",
        x ="Frequency", y = "Feature") +
  theme_minimal()
```

## Comparing subgroups

```{r, warning=FALSE}
df <- df %>%
  mutate(session_written = ifelse(session == 115, "session_115", 
                             ifelse(session == 116, "session_116",
                             ifelse(session == 117, "session_117", NA))))

ytfreq <- dfmat %>% 
  textstat_frequency(groups = df$session_written) %>%
  pivot_wider(id_cols=feature, names_from=group, values_from=frequency)
```

```{r}
#Comparing sessions 115 and 116

#Plot #1: differences in frequency by session
ggplot(ytfreq, aes(x= session_116, y= session_115)) + 
  geom_point() + theme_minimal()

#Plot #2: coloring changes by session
ytfreq$change_115_116 <- log(ytfreq$session_116 / ytfreq$session_115)
max_change <- max(abs(ytfreq$change_115_116), na.rm=TRUE)

ggplot(ytfreq, aes(x=session_116, y=session_115, fill=change_115_116)) + 
  geom_point(color="grey", shape=21) +
  scale_fill_gradientn(
    colors = c("#4575b4","white","#d73027"),
    values = scales::rescale(c(max_change*-1,0,max_change)), limits = c(max_change*-1,max_change))+
theme_bw()

#Plot 3: labeling words with biggest change
labels <- ytfreq %>% 
  rowwise() %>% 
  mutate(max_value = max(session_116, session_115)) %>% 
  filter((abs(change_115_116)>0.4 & max_value>2.5) )

ggplot(ytfreq, aes(x=session_116, y=session_115, fill=change_115_116)) + 
  geom_point(color="grey", shape=21) +
  scale_fill_gradientn(
    colors = c("#4575b4","white","#d73027"),
    values = scales::rescale(c(max_change*-1,0,max_change)), limits = c(max_change*-1,max_change)) + theme_bw() +
  geom_label_repel(data=labels, aes(label=feature), min.segment.length = 0)
```

```{r}
#Comparing sessions 116 and 117
ggplot(ytfreq, aes(x= session_117, y= session_116)) + 
  geom_point() + theme_minimal()

#Plot #2: coloring changes by session
ytfreq$change_116_117 <- log(ytfreq$session_117 / ytfreq$session_116)
max_change <- max(abs(ytfreq$change_116_117), na.rm=TRUE)

ggplot(ytfreq, aes(x=session_117, y=session_116, fill=change_116_117)) + 
  geom_point(color="grey", shape=21) +
  scale_fill_gradientn(
    colors = c("#4575b4","white","#d73027"),
    values = scales::rescale(c(max_change*-1,0,max_change)), limits = c(max_change*-1,max_change))+
theme_bw()

#Plot 3: labeling words with biggest change
labels <- ytfreq %>% 
  rowwise() %>% 
  mutate(max_value = max(session_117, session_116)) %>% 
  filter((abs(change_116_117)>0.4 & max_value>2.5) )

ggplot(ytfreq, aes(x=session_117, y=session_116, fill=change_116_117)) + 
  geom_point(color="grey", shape=21) +
  scale_fill_gradientn(
    colors = c("#4575b4","white","#d73027"),
    values = scales::rescale(c(max_change*-1,0,max_change)), limits = c(max_change*-1,max_change)) + theme_bw() +
  geom_label_repel(data=labels, aes(label=feature), min.segment.length = 0)
```

```{r}
# sims <- textstat_simil(dfmat, method="cosine")
# 
# sim_df <- as.data.frame(sims, upper = TRUE)
# 
# sim_df
```

@@

@@

```{r }
#Topics for session 115 (2017 - 2018)

lda_115 <- LDA(dfmat_115, 5)

topic_words_115 <- tidy(lda_115, matrix="beta") %>% 
  group_by(topic) %>%
  slice_max(beta, n = 10) %>%
  ungroup() %>%
  arrange(topic, -beta)

topic_words_115

topic_words_115 %>%
  mutate(term = reorder_within(term, beta, topic)) %>% 
  ggplot(aes(beta, term, fill = factor(topic))) + 
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") + scale_y_reordered() + theme_minimal()
```

```{r }
#Topics for session 116 (2019 - 2020)

lda_116 <- LDA(dfmat_116, 5)

topic_words_116 <- tidy(lda_116, matrix="beta") %>% 
  group_by(topic) %>%
  slice_max(beta, n = 10) %>%
  ungroup() %>%
  arrange(topic, -beta)

topic_words_116

topic_words_116 %>%
  mutate(term = reorder_within(term, beta, topic)) %>% 
  ggplot(aes(beta, term, fill = factor(topic))) + 
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") + scale_y_reordered() + theme_minimal()
```

```{r }
#Topics for session 117 (2021 - 2022)
raw.sum = apply(dfmat_117,1,FUN=sum) #Removing non-0 rows from dfmat
dfmat_117 = dfmat_117[raw.sum != 0, ]

lda_117 <- LDA(dfmat_117, 5)

topic_words_117 <- tidy(lda_117, matrix="beta") %>% 
  group_by(topic) %>%
  slice_max(beta, n = 10) %>%
  ungroup() %>%
  arrange(topic, -beta)

topic_words_117

topic_words_117 %>%
  mutate(term = reorder_within(term, beta, topic)) %>% 
  ggplot(aes(beta, term, fill = factor(topic))) + 
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") + scale_y_reordered() + theme_minimal()
```

## Sentiments

```{r}
#Loading df with VADER compount scores

summary_sentiment <- read_csv("https://raw.githubusercontent.com/juka19/tad_assignment3/main/data/data_w_vader.csv")
```

```{r}
ggplot(summary_sentiment, aes(x = compound)) + 
  geom_histogram(aes(y=..density..), colour="black", fill="white") +
  geom_density(alpha=.1, fill="blue") +
  labs(title="Density of compound sentiment scores",
       x ="Sentiment scores", y = "Density") +
  theme_minimal()
```

```{r}
summary_sentiment$party <- ifelse(summary_sentiment$cosponsor_D_perc > 0.66, "Democrat", ifelse(summary_sentiment$cosponsor_R_perc > 0.66, "Republican", "Both"))

wide_sentiment <- summary_sentiment %>%
  group_by(party, date) %>%
  summarise(score = mean(compound)) %>% 
  pivot_wider(names_from = party, values_from = score) %>%
  select(-c("Both", "NA"))

days <- data.frame(date = seq(as.Date("2017-01-01"),as.Date("2022-12-31"),1))

daily_sentiment <- days %>% 
  left_join(wide_sentiment) %>% 
  pivot_longer(cols = -date, names_to="party", values_to="score") %>%
  group_by(party) %>% 
  arrange(date) %>%
  mutate(score7 = data.table::frollmean(score, 7))

ggplot(daily_sentiment, aes(x=date, y = score, colour=party)) +
  geom_point(aes(y=score), size=1) + 
  geom_line(aes(y=score7)) + 
  theme_minimal() + 
  geom_smooth(method = "lm")
```

```{r}
summary_sentiment$party_full <- ifelse(summary_sentiment$cosponsor_D_perc == 1.0, "Dem", ifelse(summary_sentiment$cosponsor_R_perc == 1.0, "Rep", NA))

wide_sentiment_full <- summary_sentiment %>%
  drop_na(party_full) %>%
  group_by(party_full, date) %>%
  summarise(score = mean(compound)) %>% 
  pivot_wider(names_from = party_full, values_from = score)

days_115 <- data.frame(date = seq(as.Date("2017-01-01"),as.Date("2018-12-31"),1))

daily_sentiment_115 <- days_115 %>% 
  left_join(wide_sentiment_full) %>% 
  pivot_longer(cols = -date, names_to="party_full", values_to="score") %>%
  group_by(party_full) %>% 
  arrange(date) %>%
  mutate(score7 = data.table::frollmean(score, 7))

daily_sentiment %>% head()

ggplot(daily_sentiment_115, aes(x=date, y = score, colour=party_full)) +
  geom_point(aes(y=score), size=1) + 
  geom_line(aes(y=score7)) + 
  theme_minimal() + 
  geom_smooth(method = "lm")
```

```{r}
days_116 <- data.frame(date = seq(as.Date("2019-01-01"),as.Date("2020-12-31"),1))

daily_sentiment_116 <- days_116 %>% 
  left_join(wide_sentiment_full) %>% 
  pivot_longer(cols = -date, names_to="party_full", values_to="score") %>%
  group_by(party_full) %>% 
  arrange(date) %>%
  mutate(score7 = data.table::frollmean(score, 7))

daily_sentiment_116 %>% head()

ggplot(daily_sentiment_116, aes(x=date, y = score, colour=party_full)) +
  geom_point(aes(y=score), size=1) + 
  geom_line(aes(y=score7)) + 
  theme_minimal() + 
  geom_smooth(method = "lm")
```

```{r}
days_117 <- data.frame(date = seq(as.Date("2021-01-01"),as.Date("2022-12-31"),1))

daily_sentiment_117 <- days_117 %>% 
  left_join(wide_sentiment_full) %>% 
  pivot_longer(cols = -date, names_to="party_full", values_to="score") %>%
  group_by(party_full) %>% 
  arrange(date) %>%
  mutate(score7 = data.table::frollmean(score, 7))

ggplot(daily_sentiment_117, aes(x=date, y = score, colour=party_full)) +
  geom_point(aes(y=score), size=1) + 
  geom_line(aes(y=score7)) + 
  theme_minimal() + 
  geom_smooth(method = "lm")
```

## Some machine learning

```{r}
#Prediction model to see if can predict if bill is Republican or Democrat

#Keep only bills that are over 66% Democrat or Republican
df1 <- df %>%
  filter(party != "Both")

df1 <- df1[sample(nrow(df1)),]
df1$env <- 0
df1$env[df1$party == "Republican"] <- 1
df1$env <- factor(df1$env)

df_split <- initial_split(df1, prop=0.8) 
train_data <- training(df_split) 
test_data <- testing(df_split)

rec <- recipe(env ~ summary, data = train_data) %>% 
  step_tokenize(summary) %>% 
  step_tokenfilter(summary, max_tokens = 1e3) %>% 
  step_tfidf(summary)

model <- svm_linear(mode="classification")

wf <- workflow() %>% 
  add_recipe(rec) %>% 
  add_model(model)

model_fit <- wf %>% 
  fit(train_data)
```

```{r}
test_data$prediction <- predict(model_fit, test_data)$.pred_class

scorer <- metric_set( yardstick::accuracy, 
                      yardstick::precision, 
                      yardstick::recall, 
                      yardstick::f_meas
)

scorer(test_data, truth=env, estimate=prediction, event_level="second")
```

```{r}
#Prediction model to see if can predict if bill is focusing on health (as policy area)

df2 <- df[sample(nrow(df)),]
df2$env <- 0
df2$env[df2$policy_area == "Health"] <- 1
df2$env <- factor(df2$env)

df_split <- initial_split(df2, prop=0.8) 
train_data <- training(df_split) 
test_data <- testing(df_split)

rec <- recipe(env ~ summary, data = train_data) %>% 
  step_tokenize(summary) %>% 
  step_tokenfilter(summary, max_tokens = 1e3) %>% 
  step_tfidf(summary)

model <- svm_linear(mode="classification")

wf <- workflow() %>% 
  add_recipe(rec) %>% 
  add_model(model)

model_fit <- wf %>% 
  fit(train_data)
```

```{r}
test_data$prediction <- predict(model_fit, test_data)$.pred_class

scorer <- metric_set( yardstick::accuracy, 
                      yardstick::precision, 
                      yardstick::recall, 
                      yardstick::f_meas
)

scorer(test_data, truth=env, estimate=prediction, event_level="second")
```

>>>>

---
title: "draft_final_assignment"
author: "Andrew Wells"
date: "11/18/2022"
output: 
  html_document:
    toc: TRUE
    theme: cosmo
    toc_float: TRUE
---

# Research question

@@

-   While one might expect that the post 2016 and 2020 congresses will vary in their policies. However, it is interesting to also focus on the period between 2018 and 2020: When both chambers had different majorities and needed to cooperate.

@@

## Scraping and cleaning the data

We scraped our data from: https://data.gov/developers/apis/index.html

```{r }
df <- read_csv("https://raw.githubusercontent.com/juka19/tad_assignment3/main/data/data_11_28.csv")
head(df, 5)
```

@@

## Summary statistics

## Density of cosponsors
```{r}
ggplot(df, aes(x = cosponsor_D_perc)) + 
  geom_histogram(aes(y=..density..), colour="black", fill="white") +
  geom_density(alpha=.1, fill="blue") +
  labs(title="Density of bill cosposor party",
       x ="Cosponsor party composition", y = "Density", 
       caption = "Numbers represent proportion of cosponsors from Democratic party, 
       so 0.0 represents bills that were fully Republican and 1.0 represents 
       bills that were fully Democrat.") +
  theme_minimal()
```

@@

## Creatung a dfm from the corpus
```{r }
dfmat <- corp %>%
  tokens(remove_punct = TRUE, remove_numbers = TRUE, remove_symbols = TRUE) %>%
  tokens_remove(patter = stopwords("en")) %>%
  tokens_replace(pattern = lexicon::hash_lemmas$token, replacement = lexicon::hash_lemmas$lemma) %>%
  tokens_wordstem() %>%
  tokens_remove(c("sec","bill","act", "section", "funds", "shall","must", "use", "author","fund","provid","program","requir","divis","titl","appropri","specifi")) %>%
  dfm()

@@

@@

@@

##Sentiment analysis

```{r, warning = FALSE}
summary_sentiment <- read_csv("https://raw.githubusercontent.com/juka19/tad_assignment3/main/data/data_w_vader.csv")
summary_sentiment$party <- ifelse(summary_sentiment$cosponsor_D_perc > 0.66, "Democrat", ifelse(summary_sentiment$cosponsor_R_perc > 0.66, "Republican", "Both"))

wide_sentiment <- summary_sentiment %>%
  group_by(party, date) %>%
  summarise(score = mean(compound)) %>% 
  pivot_wider(names_from = party, values_from = score) %>%
  select(-c("Both", "NA"))
days <- data.frame(date = seq(as.Date("2017-01-01"),as.Date("2022-12-31"),1))
daily_sentiment <- days %>% 
  left_join(wide_sentiment) %>% 
  pivot_longer(cols = -date, names_to="party", values_to="score")
p4 <- ggplot(daily_sentiment, aes(x=date, y = score, colour=party)) +
  geom_point(aes(y=score), size=1) + 
  theme_minimal() + 
  geom_smooth(method = "loess", se = FALSE)+
  scale_color_manual(values = c("blue","red"))
ggplotly(p4)
```

@@

## Conclusion
